Powering Hidden Markov Model by Generative Models

### 2019-03-05 manuscript/main.tex
adding pseudo algorithms in manuscript

@@

The hypothesis of Hidden Markov Model (HMM): $\Hh := \{\bm{H} | \{\Ss, \bm{q}, A, p(\ubar{\bm{x}}|\ubar{\bm{s}}; \bm{\Phi})\}$,
\begin{itemize}
\item $\Ss$ is the set of states of HMM $\bm{H}$;
\item $\bm{q} = \left[ q_1, q_2, \cdots, q_{|\Ss|}\right]^\intercal$ initial distribution of HMM $\bm{H}$ with $|\Ss|$ is cardinality of $\Ss$, $q_k = p(s=k)$ for random state variable $s$.
\item $A$ is the transition matrix for the HMM $\bm{H}$ of size $|\Ss| \times |\Ss|$.
\item Observable signal density $p(\ubar{\bm{x}}|\ubar{\bm{s}};\bm{\Phi})$ given hidden state sequence, where $\bm{\Phi}$ is the parameter set that defines this conditional probabilistic model.
\end{itemize}

@@

Since model $\bm{H}$ contains hidden sequential variable $\ubm{s}$, we can not directly solve the maximum likelihood problem in $\autoref{eq:ml-of-hmm}$. We use expectation maximization (EM) to address the hidden variable problem by
\begin{itemize}
\item E-step: % the posterior probability of $\ubm{s}$:
  % \begin{equation}
  %   p(\ubm{s}|\ubm{x})
  % \end{equation}
  The ``expected likelihood'' function:
  \begin{equation}\label{eq:em-q-funciton}
    \Qq(\bm{H}; \bm{H}^{\mathrm{old}}) = \EE_{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})}\left[ \sum_{r=1}^{R}\log\,p(\ubmr{x}{r}, \ubmr{s}{r}; \bm{H}) \right]
  \end{equation}
\item M-step: the optimization step:
  \begin{equation}\label{eq:em-m-opt}
    \umax{\bm{H}} \Qq(\bm{H}; \bm{H}^{\mathrm{old}})
  \end{equation}
\end{itemize}

The \autoref{eq:em-m-opt} can be reformulated as:
\begin{equation}\label{eq:m-step-subs}
  \umax{\bm{H}} \Qq(\bm{H}; \bm{H}^{\mathrm{old}}) = \umax{\bm{q}}\Qq(\bm{q}; \bm{H}^{\mathrm{old}}) + \umax{{A}}\Qq({A}; \bm{H}^{\mathrm{old}}) + \umax{\bm{\Phi}}\Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}})
\end{equation}
where
\begin{align}
  \Qq(\bm{q}; \bm{H}^{\mathrm{old}}) &= \sum_{r=1}^{R} \EE_{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \left[\log\,p(\smtr{s}{1}{r}; \bm{q})\right] \label{eq:init-distribution-update}\\
  \Qq({A}; \bm{H}^{\mathrm{old}}) &= \sum_{r=1}^{R} \EE_{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \left[\log\,\sum_{t=1}^{T^{(r)}-1}p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; {A})\right] \label{eq:transition-update}\\
  \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) &= \sum_{r=1}^{R} \EE_{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \left[\log\,p(\ubmr{x}{r} | \ubmr{s}{r}; \bm{\Phi})\right]\label{eq:generative-model-update}
\end{align}

@@

@@

Since $p(\smtr{s}{1}{r}; \bm{H})$ is the probability of initial state of HMM $\bm{H}$ for $r$-th sequential, actually $q_i = p(\smtr{s}{1}{r} =i;\bm{H} ) $ for $i= 1, 2, \cdots, |\Ss|$. Solution to problem:
\begin{align}
  \bm{q}^{\mathrm{new}} &= \uargmax{\bm{q}} \Qq(\bm{q}; \bm{H}^{\mathrm{old}}), \nonumber \\
                        &\mathrm{s.t.} \sum_{i=1}^{ |\Ss| }q_i = 1 \nonumber\\
  &q_i \geq 0, \forall s.
\end{align}
is
\begin{equation}
  q_i = \frac{1}{R} \sum_{r=1}^{R} p(\smtr{s}{1}{r}=i | \ubmr{x}{r}; \bm{H}^{\mathrm{old}}), \forall\; i = 1, 2, \cdots, |\Ss|.
\end{equation}

\subsection{Transition Probability Update}
\autoref{eq:transition-update} can be written as
\begin{align}
  \Qq({A}; \bm{H}^{\mathrm{old}})
  &= \sum_{r=1}^{R} \EE_{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \left[\log\,\sum_{t=1}^{T^{(r)}-1}p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; {A})\right] \nonumber\\
  &= \sum_{r=1}^{R} \sum_{\ubmr{s}{r}}{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,\sum_{t=1}^{T^{(r)}-1}p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; {A}) \nonumber \\
  &= \sum_{r=1}^{R} \sum_{t=1}^{T^{(r)}-1} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|}\sum_{\smtr{s}{t+1}{r}=1}^{|\Ss|}{p(\smtr{s}{t}{r}, \smtr{s}{t+1}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; {A})
\end{align}

Since $A_{i, j}  = p(\smtr{s}{t+1}{r}=j|\smtr{s}{t}{r}=i; {A})$ where $A_{i, j}$ is the element of transition matrix $A$, the solution to problem:
\begin{align}
  {A}^{\mathrm{new}} = &\uargmax{{A}} \Qq({A}; \bm{H}^{\mathrm{old}}), \nonumber \\
  \mathrm{s.t.} &\hspace{0.2cm} A \cdot \bm{1} = \bm{1} \nonumber \\
                       &  A^{\intercal} \cdot \bm{1} = \bm{1} \nonumber \\
  & A_{i,j} \geq 0.
\end{align}
is
\begin{equation}
  A_{i,j}^{\mathrm{new}} = \frac{\bar{\xi}_{i,j}}{\sum_{k = 1}^{|\Ss|} \bar{\xi}_{i,k}},
\end{equation}
where
\begin{equation}
  \bar{\xi}_{i,j} = \sum_{r= 1}^{R} \sum_{t= 1}^{T^{(r)}-1}{p(\smtr{s}{t}{r}=i, \smtr{s}{t+1}{r}=j| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})}
\end{equation}

@@

\subsubsection{Generator Mixed HMM (GenM-HMM)}
\begin{figure}[!h]
  \centering
  \begin{tikzpicture}
    \tikzstyle{enode} = [thick, draw=blue, ellipse, inner sep = 1pt,  align=center]
    % \tikzstyle{nnode} = [thick, rectangle, rounded corners = 1pt,minimum size = 1.2cm,draw,inner sep = 2pt]
    \tikzstyle{nnode} = [thick, rectangle, rounded corners = 2pt,minimum size = 0.8cm,draw,inner sep = 2pt]
    \node[enode] (z1) at (-1.2,1.8) {$\bm{z}\sim p_1(\bm{z})$};
    \node[nnode] (g1) at (1,1.8) {$\bm{g}_1$};

    \node[enode] (z2) at (-1.2,0.5){$\bm{z}\sim p_{2}(\bm{z})$};
    \node[nnode] (g2) at (1,0.5) {$\bm{g}_2$};

    \node[enode] (zK) at (-1.2,-1.8) {$\bm{z}\sim p_{|\Ss|}(\bm{z})$};
    \node[nnode] (gs) at (1, -1.8) {$\bm{g}_{|\Ss|}$};

    \draw[dotted,line width=2pt] (0,-0.3) -- (0,-1.2);
    \filldraw[->] (1.9, 0.5)circle (2pt) --  (x) ;
    \draw[->] (z1) -- (g1);
    \draw[->] (g1) -- (1.8, 1.8);

@@

    % \draw[->] (z1) [in=180,out=0]  to node[right]{$\pi_1$} (g);
    % \draw[->] (z2) [in=180,out=0]  to node[above]{$\pi_2$} (g);
    % \draw[->,dashed] (zK) [in=180,out=0]  to node[right]{$\pi_K$} (g);
    %\draw[->] (g) to (x);
    \begin{scope}[xshift=0.5cm, thick, every node/.style={sloped,allow upside down}]
      \node[nnode] (m) at (3.5,-2) {Memory};
      \node[nnode] (a) at (3.5,-0.5) {$A$};

  \end{tikzpicture}
  \caption{GenM-HMM Model defined by $\bm{H} = \{\Ss, \bm{q}, A, p({\bm{x}}|{\bm{s}}; \bm{\Phi}) \}$}
  \vspace{0.1cm}
\end{figure}

For this proposal, we seek to use a generator mixed HMM scheme, termed as GenM-HMM. We define a set of generators for GenM-HMM:
\begin{equation}
  \{\bm{g}_{s}| s \in \Ss, \bm{g}_s: \bm{z}\rightarrow \bm{x}, \bm{z}\sim p_{s}(\bm{z})\}.
\end{equation}
Thus there are total $|\Ss|$ generators. $p(\bm{x}|{s}; \bm{\Phi})$ is induced as $\bm{g}_s(\bm{z}) \sim p(\bm{x}|{s}; \bm{\Phi})$ where $\bm{z} \sim p_s(\bm{z})$ for $s \in \Ss$. Let us denote the inverse of $\bm{g}_s$ as $\bm{f}_s = \bm{g}_s^{-1}$. We have the $s$-th component of the GenM-HMM model as
\begin{align}
  p(\bm{x} | s; \bm{\Phi})
  &= p_s(\bm{z}) \bigg| \det \left( \pd{\bm{z}}{\bm{x}} \right) \bigg| \nonumber \\
  &= p_s(\bm{f}_s(\bm{x})) \bigg| \det \left( \pd{\bm{f}_s(\bm{x})}{\bm{x}} \right) \bigg|
\end{align}
where $p_s(\bm{z})$ is the latent source distribution for $s=1, 2, \cdots, |\Ss|$.

Let us denote the parameter set that defines latent distribution $p_s(z)$ by $\bm{\omega}_s$ and the parameter set that defines generator $\bm{g}_s$ by $\bm{\theta}_s$. Then $\bm{\Phi} = \left\{ \bm{\theta}_s,
  \bm{\omega}_s, \forall s\in \Ss\right\}$. The problem in \autoref{eq:sub-gm} can be reformulated as:
\begin{align}
  &\umax{\bm{\Phi}} \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) \nonumber \\
  =&\umax{ \bm{\theta}_s, \bm{\omega}_s, \forall s\in \Ss} \sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}-1} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|}  p(\smtr{s}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) \left[\log\,p_{\smtr{s}{t}{r}}(\bm{f}_{\smtr{s}{t}{r}}(\bmtr{x}{t}{r})) + \log\,\bigg| \det\, \left( \pd{\bm{f}_{\smtr{s}{t}{r}}(\bmtr{x}{t}{r})}{\bmtr{x}{t}{r}} \right) \bigg| \right].
\end{align}

The diagram of GenM-HMM is shown as follows.

\subsubsection{Latent-source Mixed HMM (LatM-HMM)}
\begin{figure}[!h]
  \centering
  \begin{tikzpicture}
    \tikzstyle{enode} = [thick, draw=blue, ellipse, inner sep = 1pt,  align=center]
    % \tikzstyle{nnode} = [thick, rectangle, rounded corners = 1pt,minimum size = 1.2cm,draw,inner sep = 2pt]
    \tikzstyle{nnode} = [thick, rectangle, rounded corners = 2pt,minimum size = 0.8cm,draw,inner sep = 2pt]
    \node[enode] (z1) at (0,1.8) {$\bm{z}\sim p_1(\bm{z})$};
    \node[enode] (z2) at (0,0.5){$\bm{z}\sim p_{2}(\bm{z})$};
    \node[enode] (zK) at (0,-1.8) {$\bm{z}\sim p_{|\Ss|}(\bm{z})$}; 
    \node[enode] (x) at (5.8,1){$\ubm{x}\sim p(\ubm{x};\bm{H})$};
    % \node at (3,-1){$p(\bm{z}) = \sum_k\pi_kp_k(\bm{z})$};
    \node[nnode] (g) at (3.5,1) {$\bm{g}$};

    \node[nnode] (m) at (3.5,-2) {Memory};
    \node[nnode] (a) at (3.5,-0.5) {$A$};

>>>>

\RequirePackage{algorithm}
\RequirePackage{algorithmic}

% Attempt to make hyperref and algorithmic work together better:
% \newcommand{\theHalgorithm}{\arabic{algorithm}}

@@

The hypothesis of Hidden Markov Model (HMM): $\Hh := \{\bm{H} | \{\Ss, \bm{q}, A, p(\bm{x}|{s}; \bm{\Phi})\}$,
\begin{itemize}
\item $\Ss$ is the set of states of HMM $\bm{H}$;
\item $\bm{q} = \left[ q_1, q_2, \cdots, q_{|\Ss|}\right]^\intercal$ initial distribution of HMM $\bm{H}$ with $|\Ss|$ is cardinality of $\Ss$, $q_k = p(s=k)$ for random state variable $s$.
\item $A$ is the transition matrix for the HMM $\bm{H}$ of size $|\Ss| \times |\Ss|$.
\item Observable signal density $p({\bm{x}}|{s};\bm{\Phi})$ given hidden state sequence, where $\bm{\Phi}$ is the parameter set that defines this conditional probabilistic model.
\end{itemize}

@@

\begin{figure}[!h]
  \centering
  \begin{tikzpicture}
    \tikzstyle{enode} = [thick, draw=blue, ellipse, inner sep = 1pt,  align=center]
    % \tikzstyle{nnode} = [thick, rectangle, rounded corners = 1pt,minimum size = 1.2cm,draw,inner sep = 2pt]
    \tikzstyle{nnode} = [thick, rectangle, rounded corners = 2pt,minimum size = 0.8cm,draw,inner sep = 2pt]
    % \node[enode] (z1) at (-1.2,1.8) {Source 1}; 
    \node[nnode] (g1) at (-0.5,1.8) {Source 1};

    % \node[enode] (z2) at (-1.2,0.5){$\bm{z}\sim p_{2}(\bm{z})$};
    \node[nnode] (g2) at (-0.5,0.5) {Source 2};

    % \node[enode] (zK) at (-1.2,-1.8) {$\bm{z}\sim p_{|\Ss|}(\bm{z})$};
    \node[nnode] (gs) at (-0.5, -1.8) {Source $|\Ss|$};

    \draw[dotted,line width=2pt] (0,-0.3) -- (0,-1.2);
    \filldraw[->] (1.9, 0.5)circle (2pt) --  (x) ;

    \draw[->] (g1) -- (1.8, 1.8);

    \draw[->] (g2) -- (1.8, 0.5);

    % \draw[->] (1,-0.5) -- (1.8, -0.5);

    \draw[->] (gs) -- (1.8, -1.8);

    % \draw[->] (z1) [in=180,out=0]  to node[right]{$\pi_1$} (g);
    % \draw[->] (z2) [in=180,out=0]  to node[above]{$\pi_2$} (g);
    % \draw[->,dashed] (zK) [in=180,out=0]  to node[right]{$\pi_K$} (g);
    % \draw[->] (g) to (x);
    \begin{scope}[xshift=0.5cm, thick, every node/.style={sloped,allow upside down}]
      \node[nnode] (m) at (3.5,-2) {Memory};
      \node[nnode] (a) at (3.5,-0.5) {$\bm{A}$};

  \end{tikzpicture}
  \caption{HMM Model defined by $\bm{H} = \{\Ss, \bm{q}, \bm{A}, p({\bm{x}}|{\bm{s}}; \bm{\Phi}) \}$}
  \vspace{0.1cm}
\end{figure}

Since model $\bm{H}$ contains hidden sequential variable $\ubm{s}$, we can not directly solve the maximum likelihood problem in \autoref{eq:ml-of-hmm}. We use expectation maximization (EM) to address the hidden variable problem by
\begin{itemize}
\item E-step: % the posterior probability of $\ubm{s}$:
  % \begin{equation}
  %   p(\ubm{s}|\ubm{x})
  % \end{equation}
  The ``expected likelihood'' function:
  \begin{equation}\label{eq:em-q-funciton}
    \Qq(\bm{H}; \bm{H}^{\mathrm{old}}) = \EE_{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})}\left[ \sum_{r=1}^{R}\log\,p(\ubmr{x}{r}, \ubmr{s}{r}; \bm{H}) \right]
  \end{equation}
\item M-step: the optimization step:
  \begin{equation}\label{eq:em-m-opt}
    \umax{\bm{H}} \Qq(\bm{H}; \bm{H}^{\mathrm{old}})
  \end{equation}
\end{itemize}

The \autoref{eq:em-m-opt} can be reformulated as:
\begin{equation}\label{eq:m-step-subs}
  \umax{\bm{H}} \Qq(\bm{H}; \bm{H}^{\mathrm{old}}) = \umax{\bm{q}}\Qq(\bm{q}; \bm{H}^{\mathrm{old}}) + \umax{{A}}\Qq({A}; \bm{H}^{\mathrm{old}}) + \umax{\bm{\Phi}}\Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}})
\end{equation}
where
\begin{align}
  \Qq(\bm{q}; \bm{H}^{\mathrm{old}}) &= \sum_{r=1}^{R} \EE_{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \left[\log\,p(\smtr{s}{1}{r}; \bm{q})\right] \label{eq:init-distribution-update}\\
  \Qq(\bm{A}; \bm{H}^{\mathrm{old}}) &= \sum_{r=1}^{R} \EE_{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \left[\log\,\sum_{t=1}^{T^{(r)}-1}p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; {A})\right] \label{eq:transition-update}\\
  \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) &= \sum_{r=1}^{R} \EE_{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \left[\log\,p(\ubmr{x}{r} | \ubmr{s}{r}; \bm{\Phi})\right]\label{eq:generative-model-update}
\end{align}

@@

We summarize the optimization algorithm as:

\begin{algorithm}[H]
  \caption{Meta algorithm for HMM powered by Generative Models}
  \begin{algorithmic}[1]
    \STATE {\bfseries Input:}{
      Building $\bm{H}^{\mathrm{old}}, \bm{H} \in \Hh$ gives: \\
      $\bm{H}^{\mathrm{old}} = \{\Ss, \bm{q}^{\mathrm{old}}, A^{\mathrm{old}}, p(\bm{x}|s; \bm{\Phi}^{\mathrm{old}})\}$, $\bm{H} = \{\Ss, \bm{q}, A, p(\bm{x}|s; \bm{\Phi})\}$
    }, \\
    \STATE Initialize $\bm{H}$
    \STATE $\bm{H}^{\mathrm{old}} \gets \bm{H}$
    \FOR { $\bm{H}$ not converge}
    \STATE Sample a batch of data $\left\{ \ubmr{x}{r} \right\}_{r=1}^{R_b}$ from the dataset $\hat{p}(\ubm{x})$
    \STATE Compute $p(\smtr{s}{t}{r} | \ubmr{x}{r}; \bm{H}^{\mathrm{old}})$, $p(\smtr{s}{t}{r}, \smtr{s}{t+1}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})$ by forward/backward algorithm;
    \STATE $\bm{q} \gets \uargmin{\bm{q}}\, \Qq(\bm{q}; \bm{H}^{\mathrm{old}})$ by \autoref{eq:update-initial-state-prob};
    \STATE $\bm{A} \gets \uargmin{\bm{A}}\Qq(\bm{A}; \bm{H}^{\mathrm{old}})$ by \autoref{eq:update-transition-prob};
    \STATE $\bm{\Phi} \gets \uargmin{\bm{\Phi}}\Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}})$ by calling Algorithm \autoref{algo:opt-Phi-GenHMM} or \autoref{algo:opt-Phi-LatHMM};
    \STATE $\bm{H}^{\mathrm{old}} \gets \bm{H}$

    \ENDFOR
  \end{algorithmic}
\end{algorithm}

@@

Since $p(\smtr{s}{1}{r}; \bm{H})$ is the probability of initial state of HMM $\bm{H}$ for $r$-th sequential, actually $q_i = p(\smtr{s}{1}{r} =i;\bm{H} ) $ for $i= 1, 2, \cdots, |\Ss|$. Solution to problem:
\begin{align}
  \bm{q}^{\mathrm{new}} &= \uargmax{\bm{q}} \Qq(\bm{q}; \bm{H}^{\mathrm{old}}), \nonumber \\
                        &\mathrm{s.t.} \sum_{i=1}^{ |\Ss| }q_i = 1 \nonumber\\
                        &q_i \geq 0, \forall s.
\end{align}
is
\begin{equation}\label{eq:update-initial-state-prob}
  q_i = \frac{1}{R} \sum_{r=1}^{R} p(\smtr{s}{1}{r}=i | \ubmr{x}{r}; \bm{H}^{\mathrm{old}}), \forall\; i = 1, 2, \cdots, |\Ss|.
\end{equation}

\subsection{Transition Probability Update}
\autoref{eq:transition-update} can be written as
\begin{align}
  \Qq(\bm{A}; \bm{H}^{\mathrm{old}})
  &= \sum_{r=1}^{R} \EE_{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \left[\log\,\sum_{t=1}^{T^{(r)}-1}p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; {A})\right] \nonumber\\
  &= \sum_{r=1}^{R} \sum_{\ubmr{s}{r}}{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,\sum_{t=1}^{T^{(r)}-1}p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; {A}) \nonumber \\
  &= \sum_{r=1}^{R} \sum_{t=1}^{T^{(r)}-1} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|}\sum_{\smtr{s}{t+1}{r}=1}^{|\Ss|}{p(\smtr{s}{t}{r}, \smtr{s}{t+1}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; {A})
\end{align}

Since $\bm{A}_{i, j}  = p(\smtr{s}{t+1}{r}=j|\smtr{s}{t}{r}=i; {A})$ where $A_{i, j}$ is the element of transition matrix $A$, the solution to problem:
\begin{align}\label{eq:update-transition-prob}
  \bm{A}^{\mathrm{new}} = &\uargmax{\bm{A}} \Qq(\bm{A}; \bm{H}^{\mathrm{old}}), \nonumber \\
  \mathrm{s.t.} &\hspace{0.2cm} \bm{A} \cdot \bm{1} = \bm{1} \nonumber \\
                          &  \bm{A}^{\intercal} \cdot \bm{1} = \bm{1} \nonumber \\
                          & \bm{A}_{i,j} \geq 0.
\end{align}
is
\begin{equation}
  \bm{A}_{i,j}^{\mathrm{new}} = \frac{\bar{\xi}_{i,j}}{\sum_{k = 1}^{|\Ss|} \bar{\xi}_{i,k}},
\end{equation}
where
\begin{equation}
  \bar{\xi}_{i,j} = \sum_{r= 1}^{R} \sum_{t= 1}^{T^{(r)}-1}{p(\smtr{s}{t}{r}=i, \smtr{s}{t+1}{r}=j| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})}
\end{equation}

@@

\subsubsection{Generator Mixed HMM (GenHMM)}\label{subsubsec:GenHMM}
\begin{figure}[!h]
  \centering
  \begin{tikzpicture}
    % \tikzstyle{enode} = [thick, draw=blue, circle, inner sep = 3pt,
    % align=center]
    \tikzstyle{enode} = [thick, draw=blue, ellipse, inner sep = 2pt,  align=center]
    \tikzstyle{nnode} = [thick, rectangle, rounded corners = 2pt,minimum size = 0.8cm,draw,inner sep = 2pt]

    \node[enode] (z1) at (-1.2,1.8) {$\bm{z}\sim p_{s,1}(\bm{z})$};
    \node[nnode] (g1) at (1,1.8) {$\bm{g}_{s,1}$};

    \node[enode] (z2) at (-1.2,0.5){$\bm{z}\sim p_{s,2}(\bm{z})$};
    \node[nnode] (g2) at (1,0.5) {$\bm{g}_{s,2}$};

    \node[enode] (zK) at (-1.2,-1.8) {$\bm{z}\sim p_{s,K}(\bm{z})$};
    \node[nnode] (gs) at (1, -1.8) {$\bm{g}_{s,K}$};

    \node[enode] (x) at (4.5,0){$\bm{x}\sim p(\bm{x}| s; \bm{\Phi}_{s})$};
    % \node at (3,-1){$p(\bm{z}) = \sum_k\pi_kp_k(\bm{z})$};

    \draw[dotted,line width=2pt] (0,-0.3) -- (0,-1.2);
    \filldraw[->] (1.9, 0.5)circle (2pt) --  node[above=0.2]{${\kappa}\sim \bm{\pi}_{s}$} (x)  ;
    \draw[->] (z1) -- (g1);
    \draw[->] (g1) -- (1.8, 1.8);

@@

    % \node[enode] (z) at (0,0) {$\bm{z}\sim p(\bm{z})$};
    % \node[enode] (x) at (5.5,0){$\bm{x}\sim p(\bm{x}; \bm{\Phi})$};
    %               % \node at (5.2,-1) {$p(\bm{x};\bm{\Phi}) = \textstyle\sum_{k=1}^K \pi_k  p_k(\bm{x})$};
    % \node[nnode] (g1) at (2.6,1.8) {$\bm{g}_1$};
    % \node[nnode] (g2) at (2.6,0.5) {$\bm{g}_2$};
    % \node[nnode] (gk) at (2.6,-1.8) {$\bm{g}_K$};
    % \draw[dotted,line width=2pt] (2.6,-0.3) -- (2.6,-1.2);
    % \draw[->] (z) [in= 180, out =0] to (g1);
    % \draw[->] (z) [in= 180, out =0] to (g2);
    % \draw[->] (z) [in= 180, out =0] to (gk);
    % \filldraw[->] (3.7, 0.5)circle (2pt) -- node[above=0.2]{$\bm{s}\sim \bm{\pi}$} (x) ;
    %               % \draw[->] (3,-0.8) -- (3.5, -0.8);
    % \draw[->] (g1) -- (3.5,1.8);
    % \draw[->] (g2) -- (3.5, 0.5);
    % \draw[->] (gk) -- (3.5, -1.8);
  \end{tikzpicture}
  \caption{Source $s$ powered by generator mixed generative model (GenHMM).}
  \vspace{0.1cm}
\end{figure}

Base on the generative model assumption:

\begin{equation}
  p(\bm{x}| s; \bm{\Phi}_{s}) = \sum_{\kappa=1}^{K}\pi_{s, \kappa} p(\bm{x}| s, \kappa; \bm{\Phi}_{s, \kappa})
\end{equation}
where $\bm{\Phi}_s = \left\{ \bm{\Phi}_{s, \kappa}| \kappa = 1, 2, \cdots, K. \right\}$, and 
\begin{equation}
  \sum_{\kappa = 1}^{K} \pi_{s, \kappa}= 1
\end{equation}

% Let us denote the inverse of $\bm{g}_s$ as $\bm{f}_s = \bm{g}_s^{-1}$.
% For this case we assume the latent sources are Gaussian. We write the emission probability model as a $K$ generator mixed model:
% \begin{align}
    %     p(\bm{x} | s; \bm{\Phi}) &= \sum_{\kappa = 1}^{K} \pi_{s, \kappa} \Nn\left( \bm{f}_{s,\kappa}(\bm{x}); \bm{\mu}_{s,\kappa}, \bm{C}_{s,\kappa}  \right)\bigg|\det\,\left( \pd{\bm{f_{s,\kappa}(\bm{x})}}{\bm{x}} \right)\bigg| \nonumber \\
    %     \sum_{\kappa = 1}^{K} \pi_{s, \kappa}&= 1
                                                 %   \end{align}

The help function should be revised to deal with the new latent variable $\kappa$ into:
\begin{align}\label{eq:obj-q-gen-mix}
  \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) & = \sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}-1} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|} \sum_{\smtr{\kappa}{t}{r}=1}^{K}p(\smtr{s}{t}{r}, \smtr{\kappa}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})  \log\,p(\smtr{\kappa}{t}{r}, \bmtr{x}{t}{r}| \smtr{s}{t}{r}; \bm{\Phi})\nonumber \\
  =& \sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}-1} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|}  \sum_{\smtr{\kappa}{t}{r}=1}^{K}p(\smtr{s}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})p(\smtr{\kappa}{t}{r}|\smtr{s}{t}{r}, \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) \bigg[\log\,\pi_{\smtr{s}{t}{r},\smtr{\kappa}{t}{r}} \nonumber \\
                                        & + \log\, p(\bmtr{x}{t}{r} | \smtr{s}{t}{r}, \smtr{\kappa}{t}{r}; \bm{\Phi}) \bigg]
\end{align}
                                          %                                           \Nn\left( \bm{f}_{\smtr{s}{t}{r},\smtr{\kappa}{t}{r}}(\bm{x}); \bm{\mu}_{\smtr{s}{t}{r},\smtr{\kappa}{t}{r}}, \bm{C}_{\smtr{s}{t}{r},\smtr{\kappa}{t}{r}} \right) \log\,\bigg| \det\, \left( \pd{\bm{f}_{\smtr{s}{t}{r},\smtr{\kappa}{t}{r}}(\bmtr{x}{t}{r})}{\bmtr{x}{t}{r}} \right) \bigg| \bigg].
                                          %   \end{align}
In \autoref{eq:obj-q-gen-mix}, $p(s_t| \ubm{x}, \bm{H}^{\mathrm{old}})$ is computed by forward/backward algorithm. The posterior of $\kappa$ is:
\begin{align}\label{eq:kappa-posterior}
  p(\kappa| s, \ubm{x}; \bm{H}^{\mathrm{old}})
  &=  \frac{p(\kappa, \ubm{x}| s; \bm{H}^{\mathrm{old}})}{p(\ubm{x}| s,\bm{H}^{\mathrm{old}})} \nonumber \\
  & = \frac{\pi_{s, \kappa} p(\ubm{x}| s, \kappa, \bm{H}^{\mathrm{old}})}{\sum_{\kappa=1}^{K} \pi_{s, \kappa} p(\ubm{x}| s, \kappa,\bm{H}^{\mathrm{old}})} \nonumber \\
  & = \frac{\pi_{s, \kappa} p(\bm{x}| s, \kappa, \bm{H}^{\mathrm{old}})}{\sum_{\kappa=1}^{K}  \pi_{s, \kappa} p(\bm{x}| s, \kappa,\bm{H}^{\mathrm{old}})} 
  % & = \frac{\pi_{s_t, \kappa_t} \Nn\left( \bm{f}_{{s}_{t},{\kappa}_{t}}(\bm{x}); \bm{\mu}_{{s}_{t},{\kappa}_{t}}, \bm{C}_{{s}_{t},{\kappa}_{t}} \right) \bigg| \det\, \left( \pd{\bm{f}_{{s}_{t},{\kappa}_{t}}(\bmtr{x}{t}{r})}{\bmtr{x}{t}{r}} \right) \bigg| }{\sum_{\kappa_t=1}^{K}\pi_{s_t, \kappa_t} \Nn\left( \bm{f}_{{s}_{t},{\kappa}_{t}}(\bm{x}); \bm{\mu}_{{s}_{t},{\kappa}_{t}}, \bm{C}_{{s}_{t},{\kappa}_{t}} \right) \bigg| \det\, \left( \pd{\bm{f}_{{s}_{t},{\kappa}_{t}}(\bmtr{x}{t}{r})}{\bmtr{x}{t}{r}} \right) \bigg|} \nonumber \\
\end{align}
where the last equation is due to the fact that only $\bm{x}_t$ among sequence $\ubm{x}$ depends on $s_t, \kappa_t$.

    %     \begin{align}

    %   \end{align}

The latent prior for mixture of each source $s$ is obtained by solving the following problem:
\begin{align}
  \pi_{s, \kappa} & = \uargmax{pi_{s, \kappa}} \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) \\ \nonumber
                  & s.t. \, \sum_{\kappa=1}^{K} \pi_{s, \kappa}= 1
\end{align}
which gives the solution:
\begin{equation}\label{eq:mix-latent-parameter-solution}
  \pi_{s, \kappa} = \frac{\sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}-1} p(\smtr{s}{t}{r} =s, \smtr{\kappa}{t}{r}=\kappa | \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) }{\sum_{k =1}^{K}\sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}-1} p(\smtr{s}{t}{r} =s, \smtr{\kappa}{t}{r}=k | \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) }
\end{equation}
where $p(s, \kappa | \ubm{x}; \bm{H}^{\mathrm{old}}) = p(s| \ubm{x}; \bm{H}^{\mathrm{old}}) p(\kappa | s, \ubm{x}; \bm{H}^{\mathrm{old}})$ that can be computed by results of forward/backward and \autoref{eq:kappa-posterior}.

In implementation, GenHMM uses the generator mixed emission model with $\kappa$-th component as:
